\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
First draft of full-abstraction subsection.

\subsection{Full abstraction}

@@

We are now ready to define plugging of components into contexts followed by contextual equivalence:
%The remaining defintions
\begin{definition}[Plugging a component into a context]
  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
\end{definition}
\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
  \label{def:contextual-equivalence}
  ~
  \begin{description}
  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
  \begin{equation*}
    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
  \end{equation*}
  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
  \begin{equation*}
    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
  \end{equation*}
\end{description}
\end{definition}
With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
% Full abstraction thm
\begin{theorem}
  \label{thm:full-abstraction}
  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
  \begin{gather*}
    \src{\comp_1} \sconeq \src{\comp_2}\\
    \Updownarrow\\
    \src{\comp_1} \tconeq \src{\comp_2}
  \end{gather*}
\end{theorem}
Readers unfamiliar with full-abstraction may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantee LSE and WBCF.
\todo[inline]{Write this, i.e.\ explain why full-abstractness prove LSE and WBCF.}

% What does it actually mean?
% Relies on LSE and WBCF of \srccm - which is obvious from the operational semantics (or defines what it means here?).

>>>>

\subsection{Full abstraction}
All that is left before we state the full-abstraction theorem is to define how programs are lifted to executable configurations and contextual equivalence for each of the two capability machines we have defined.

@@

Contextual equivalence roughly says that two components behave the same no matter what context we plug them into.
%The remaining defintions
\begin{definition}[Plugging a component into a context]
  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
\end{definition}
\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
  \label{def:contextual-equivalence}
  ~
  \begin{description}
  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
  \begin{equation*}
    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
  \end{equation*}
  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
  \begin{equation*}
    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
  \end{equation*}
\end{description}
where $\Phi\sterm[i]{\src{\ta,\stkb}} \text{ iff } \Phi \nstep[i]{\src{\ta,\stkb}} \halted \quad$ and $\quad \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}$
\end{definition}
With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
% Full abstraction thm
\begin{theorem}
  \label{thm:full-abstraction}
  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
  \begin{gather*}
    \src{\comp_1} \sconeq \src{\comp_2}\\
    \Updownarrow\\
    \src{\comp_1} \tconeq \src{\comp_2}
  \end{gather*}
\end{theorem}
Readers unfamiliar with fully-abstract compilation may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantees LSE and WBCF.
Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
This means that any property the source language has must somehow be there after compilation whether or not it is a property of the target language.
If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that it behaviorally matches exactly.
In our case, LSE and WBCF was built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
In order to enforce these properties, components on \trgcm{} use \stktokens{}.
Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
